Does an earthquake “know” how big it will be? A neural-net aided study

### 2024-07-05 article.tex
small changes mostly typos

To this end, we construct a neural-based model that predicts the magnitude of a given earthquake given the short and long term seismic history prior to its occurrence. Our model is trained on seismic catalogs containing information about past earthquakes, including location, time, and magnitude. In some experiments, we also incorporate focal mechanism information, demonstrating improved performance. Importantly, the model is not tasked with predicting the timing and location of the event, as they are explicitly provided. Thus, we separate out the task of modeling nucleation statistics and isolate the question of magnitude predictability. If our model performs better than a random draw from a GR distribution or its variants, as we will indeed demonstrate is the case, we assert that at least some information about the magnitude of a specific earthquake is extractable from cataloged properties alone.

@@

MAGNET is composed of two main components. First, the catalog up to time $t$ (not including) is encoded by a long-short term memory (LSTM) units, which produces a latent representation of the seismic history \cite{hochreiter_long_1997}.
Second, this representation is combined with a space-time query specifying the coordinates, $\bf{x}_i$, $t_i$, of a future earthquake and passed into a fully connected neural network (FCNN) that produces a parametrized PDF of the event's magnitude, $p_{\textbf{x}_i, t_i}(m)$, see Fig. \ref{fig:intro_fig}a.

@@

\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/intro_fig.png}
	\caption{\textbf{a} MAGnitude Neural EsTimation model (MAGNET) task and architecture. Seismic history is continuously encoded using a Long Short-Term Memory neural network. For each earthquake, the encoded history up to its occurrence time ($t_i$) is concatenated with the time and location of the earthquake (``spacetime query''). The combined information is fed into a fully-connected neural network that outputs a probability density function (PDF) of the predicted magnitude at that time and location. \textbf{b} Performance on Major Earthquakes: MAGNET's resulting PDFs for a selection of well-known major earthquakes worldwide. The red curve represents the model's predicted PDF and the dashed black curve represents the Gutenberg-Richter (GR) magnitude-frequency distribution, a common naive benchmark. For most earthquakes shown here, the likelihood of the observed magnitude (red star) is higher for MAGNET than for the GR benchmark (black circle). Notably, the Tōhoku aftershock shows a qualitatively different behavior.
    }
\label{fig:intro_fig}
\end{figure}

All metrics reported in this work were evaluated over a time span that is not included in training period. The presented metrics demonstrate that our model consistently and significantly outperforms all benchmarks across all test regions, indicating a significant information gain in forecasting earthquake magnitudes prior to their occurrence.
%This finding directly suggests that the seismic system undergoes preparatory processes tailored to the impending event's magnitude before its initiation. Examining these results across multiple catalogs and regions highlights the robustness of our methodology and the generalizability of our conclusion, paving the way for further exploration of earthquake predictability.

\section{Results} \label{sec:results}
As explained above, the model outputs a PDF $p_{\textbf{x}_i, t_i}(m_i)$ for each earthquake in the test set. We denote by $\ell_i$ the log-likelihood of the observed magnitude $m_i$ as predicted by MAGNET, 
\begin{align}
    \mathcal{L} &= -\langle \ell_i \rangle\ , 
    &
    \ell_i&=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)
    \label{eq:likelihood}
\end{align}
 $\mathcal{L}$ is the average minus log-likelihood, which the model is trained to minimize.  The values of $\ell_i$ for MAGNET and the leading benchmark are presented in Extended Material Fig. \ref{fig:labels_likelihood}.

@@

In addition to comparing the log-likelihood of the observed data, one can convert the prediction problem to a binary question: will the earthquake at the spacetime query $\textbf{x}_i, t_i$ have a higher magnitude than a threshold $m_t$? This can be easily calculated from the prediction, as the probability that $m_i>m_t$ is simply $\int_{m_t}^{\infty}p(m')dm'$.
This task, asking whether a given event will be a large earthquake, is more interpretable and can be evaluated using standard binary classification metrics such as area under receiver operator (ROC) curve \cite{Murphy} and under the interpolated precision-recall (PR) curve \cite{buttcher_information_2010} which suit better an imbalanced data set such as ours. Fig.~\ref{fig:metrics}d-f and g-i show these metrics for all three regions in examined and compare against the GR-variant models. Again, MAGNET outperforms all benchmarks.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_barplots.pdf}
    \caption{
        \textbf{Benchmarking MAGNET. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicates a better preforming model. Scores conditioned above the temporal incompleteness are marked by the horizontal dashed line on each corresponding bar. \textbf{d, e, f}, Receiver Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the existence of a large ($m>=4$) event. The performance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, color coded identically to the bar plots. For the AUC metrics, a higher score indicates a better preforming classifier.
        }
        \label{fig:metrics}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/info_over_time_hauksson.pdf}
    \caption{
        \textbf{Information gain in MAGNET. a}, Information gain of individual events in the Southern California test set. Scattered dots indicate magnitude at event index, colored by the information gain per event (scale indicated by colorbar), here warm is an advantage for MAGNET, while cool is an advantage for the benchmark. Secondary horizontal axis (grey) indicates corresponds origin time. Cumulative Information gain (CIG) (dashed blue), temporally conditioned CIG (dashed yellow) and spatially conditioned CIG (dashed red) curves are superimposed on the scatter, demonstrating a constantly increasing function. \textbf{b-d}, Information gain (dashed blue) temporally conditioned CIG (dashed yellow) and spatially conditioned CIG (dashed red) curves for Southern California, New Zealand and Japan test sets, respectively, all three regions examined in this study. An exception from the increasing trend can be observed for the spatially conditioned CIG of New Zealand (red curve in c). \textbf{e-g} Spatial distribution of spatially conditioned information gain for all regions in study. Maps show the difference in mean information content per bin, model's advantage is indicated by warm colors, disadvantage by cool colors, as defined by the colorbar adjacent to f.
    }
    \label{fig:info_gain_over_time}
\end{figure}

@@

Our analysis reveals a persistent increase in information gain across the test set for all investigated regions (Fig. \ref{fig:info_gain_over_time}b-d). Specifically, we show an average information gain of $>0.1$ nats per earthquake in the same region in Japan where Ogata et.~al.~\cite{ogata_exploring_2018} found no information gain. Similar plots for the other studied regions are shown in Extended Material Fig.~\ref{fig:nz_japan_info_gain}.

@@

An obvious such artifact is ``short-time incompleteness'' (STI): the fact that immediately following a large event the signal of subsequent small events might be buried in the coda of the mainshock~\cite{kagan_short-term_2004, stockman_forecasting_2023}. Thus, immediately after a large event it is less likely that a small event will be included in the catalog, not because it did not occur, but rather because it was not detected. Indeed, our model consistently allocates less probability mass to low magnitudes shortly after large events (see the Tohoku aftershock in Fig.~\ref{fig:intro_fig}b), and some of the information gain may be attributed to this effect.

@@

The instantaneous completeness magnitude, $m_c(t)$, is defined as the minimal magnitude above which all earhtquakes can be detected. If we choose $\tilde m \geq m_c$ then, by definition, the model cannot present an information gain due to the STI effect.

We thus take $\Tilde{m}=m_c(t)$, which we estimate dynamically using the maximum curvature method \cite{wiemer_minimum_2000} within a window of 150 past and 150 current and future events. The resulting $m_c(t)$ curves are presented in the Extended Material Fig. \ref{fig:temp_incompleteness}. The conditioned cumulative information gain (CIG) is then recalculated and presented as the yellow dashed curved in Fig.  \ref{fig:info_gain_over_time} for all three tested regions. 
Indeed, factoring out the temporal incompleteness reduces the information gain, though it is still significant: evidently, the conditioned CIG curves show a steadily increasing trend. the horizontal dashed lines on the bars in Fig \ref{fig:metrics}a-c indicate the mean conditioned scores, also presented in Table \ref{tab:mean_ll_all_benchmarks}. EM Table \ref{tab:mean_ll_benchmarks_main_text} displays the conditioned scores for all benchmarks tested in this study.

Similarly to the temporal domain, a spurious information gain may also be caused by temporally constant but spatially varying properties. For example, fault stresses, geometry and lithology may change in space resulting in spatially-dependent GR distribution \cite{amitrano_brittle-ductile_2003, scholz_stress_2015, herrmann_revealing_2022, taroni_earthquake_2023}, and the local completeness magnitudes is also generally spatially dependent. In such a case, the information gain of MAGNET does not imply that the magnitudes are correlated in time. 
To examine this, we preform two additional tests. First, we construct a spatially dependent GR distributions where $m_c$ is held constant ($=m_c^{\mbox{\footnotesize (train set)}}$). MAGNET outperforms the spatially varying GR benchmark for all regions in the unconditioned $\mathcal{L}$ score, as can be seen in EM Table \ref{tab:mean_ll_all_benchmarks}. The spatially varying GR shows an advantage over MAGNET for solely the New Zealand data set for the temporally conditioned $\mathcal{L}$. This advantage may arise from a the measurement artifact of spatial variation of $m_c$ originating from the properties of the seismic network.

In order to account for such spatially dependent artifacts we preform a simialrecond calculation. We once again condition MAGNET's results on a threshold according to Eq. \ref{eq:conditioned_likelihood}, this time choosing a spatially varying completeness $\Tilde{m}=m_c(x,y)$. We estimate $m_c(x,y)$ for a grid of coordinates spaced every $0.1^\circ$, for all events within a square centered at each grid point, with a side length of $0.2^\circ$.

@@

\section{Possible affects on operational earthquake forecasting and prediction}
Identifying precursory patterns in the seismic history can advance the effort of earthquake prediction a great deal\cite{mousavi_deep-learning_2022, mousavi_machine_2023, mignan_neural_2020, karimpouli_explainable_2023, bergen_machine_2019}. Nevertheless, discovering such signals with enough accuracy to prove useful for operational systems may still be further down the road.
However, an immediate application of the model may be a quick estimation of magnitude for early warning systems. Such a system does not require additional data sources or assumptions on the model as the timing and location of the earthquake to be estimated are known or estimated. 
Another immediate application is to incorporate such a history-dependent magnitude distribution into point-process rate forecasting models, such as the epidemic type aftershock sequence model (ETAS) or similar \cite{ogata_statistical_1988, ogata_statistics_2017, dascher-cousineau_using_2023}. As discussed above, such models currently assume no statistical dependence of the magnitude. Since larger earthquakes produce more aftershock, a more accurate estimation of magnitudes will have a positive compounding effect on the resulting statistics.

@@

Outputs of all encoders are concatenated and fed to a Fully Connected Neural Network (FCNN). The sizes of layers in the FCNN, as the number of layers themselves, are optimized by cross validation. A $tanh$ activation is used for the hidden layers, and a $\texttt{softplus}$ as an activation on the output layer.

@@

\subsection{Definition of benchmark models}
\subparagraph{Last $n$ events} The Gutenberg- Richter (GR) distribution fitted for the past $n$ events. This method follows the method presented in Gulia \& Wiemer\cite{gulia_real-time_2019} for constructing a b-value time-series. If in the last $n$ events there are less than 10 earthquakes of magnitudes above $m_c^{train}$, then more events from the past are included, until this condition is met. $m_c^{train}$ is estimated using the maximal curvature method\cite{wiemer_minimum_2000}.

@@

\subparagraph{Spatially varying GR} Estimation of the local $b$ value is done by following the method of Taroni et.~al.~\cite{taroni_highdefinition_2021}. Seismicity data is 
binned into $0.1^\circ \times 0.1^\circ$ bins, and the GR distribution is fitted for each bin. $b$ for a specified location $x_i$ is calculated by the average of all $b$ values in bins within 30km around $x_i$. Estimations which consider less than 150 events are discarded.

@@

\newpage
\section*{Acknowledgements}
We thank Yehuda Ben-Zion, Assaf Inbal, Kelian Dascher-Cousineau, David Marsan and Eugenio Lippiello, Jiancang Zhuang and Yosihiko Ogata. for thoughtful discussions. YBS is supported by ISF grant 1907/22 and by Google Gift grant.

@@

\newpage
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/detailed_architecture.pdf}
    \caption{\textbf{MAGNET architecture.} Described in detail in the methods section.}
    \label{fig:architecture}
\end{figure}

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.png}
	\caption{
            \textbf{Information gain of MAGNET over the benchmark, New Zealand and Japan data sets.} Scattered dots indicate magnitude at event index, colored by the information gain per event. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks. Low magnitude data points in a, c and d are subsampled to ease on image rendering.
         }
\label{fig:nz_japan_info_gain}
\end{figure}

\newpage
\begin{figure}[h!]
    \centering
        \includegraphics[width=1\textwidth]{figures/temporal_incompleteness.png}
    \caption{
    \textbf{Temporal Incompleteness for test sets of all regions examined.} Calculated temporal incompleteness for test sets is presented in red, superimposed over the events above (black) and below (grey) the completeness magnitude of the train set, $m_c^{(train)}$. Events are plotted by the index of their appearance in the test set. Events of low magnitude, $m_c^{(train)}$ (grey), that are not included in the evaluation, are plotted at their interpolated location along the horizontal axis. Data for Japan (c) is displayed only near the 2011 Tōhoku earthquake to reduce data density and enhance clarity of the figure. Low magnitude data points in b and c are subsampled to ease on image rendering.
    }
    \label{fig:temp_incompleteness}
\end{figure}

@@

\newpage
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/likelihood_scatter.png}
    \caption{
    \textbf{Log-likelihood of the true magnitude, $\ell_i$}. The likelihood for the true magnitude label of the test set for each data set, $\ell_i=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)$, as defined in main text. $\ell_i$ values of MAGNET and the last 300 events benchmark are shown in red and orange respectively. The blue dashed line indicates the values that the GR benchmark would obtain. Low magnitude data points in b and c are subsampled to ease on image rendering.
    }
    \label{fig:labels_likelihood}
\end{figure}

>>>>

To this end, we construct a neural-based model that predicts the magnitude of a given earthquake given the short and long term seismic history prior to its occurrence. Our model is trained on seismic catalogs containing information about past earthquakes, including location, time, and magnitude. In some experiments, we also incorporate focal mechanism information. Importantly, the model is not tasked with predicting the timing and location of the event, as they are explicitly provided. Thus, we separate out the task of modeling nucleation statistics and isolate the question of magnitude predictability. If our model performs better than a random draw from a GR distribution or its variants, as we will indeed demonstrate is the case, we assert that at least some information about the magnitude of a specific earthquake is extractable from cataloged properties alone.

@@

MAGNET is composed of two main components. First, the catalog up to time $t$ (not including) is encoded by long-short term memory (LSTM) units, which produces a latent representation of the seismic history \cite{hochreiter_long_1997}.
Second, this representation is combined with a space-time query specifying the coordinates, $\bf{x}_i$, $t_i$, of a future earthquake and passed into a fully connected neural network (FCNN) that produces a parametrized PDF of the event's magnitude, $p_{\textbf{x}_i, t_i}(m)$, see Fig. \ref{fig:intro_fig}a.

@@

\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/intro_fig.png}
	\caption{\textbf{a} MAGnitude Neural EsTimation model (MAGNET) task and architecture. Seismic history is continuously encoded using a Long Short-Term Memory neural network. For each earthquake, the encoded history up to its occurrence time ($t_i$) is concatenated with the time and location of the earthquake (``spacetime query''). The combined information is fed into a fully-connected neural network that outputs a probability density function (PDF) of the predicted magnitude at that time and location. \textbf{b} Performance on major earthquakes: MAGNET's resulting PDFs for a selection of well-known major earthquakes worldwide. The red curve represents the model's predicted PDF and the dashed black curve represents the Gutenberg-Richter (GR) magnitude-frequency distribution, a common naive benchmark. For most earthquakes shown here, the likelihood of the observed magnitude (red star) is higher for MAGNET than for the GR benchmark (black circle). Notably, the Tōhoku aftershock shows a qualitatively different behavior.
    }
\label{fig:intro_fig}
\end{figure}

All metrics reported in this work were evaluated over a time span that is not included in the training period. The presented metrics demonstrate that our model consistently and significantly outperforms all benchmarks across all test regions, indicating a significant information gain in forecasting earthquake magnitudes prior to their occurrence.
%This finding directly suggests that the seismic system undergoes preparatory processes tailored to the impending event's magnitude before its initiation. Examining these results across multiple catalogs and regions highlights the robustness of our methodology and the generalizability of our conclusion, paving the way for further exploration of earthquake predictability.

\section{Results} \label{sec:results}
As explained above, the model outputs a PDF $p_{\textbf{x}_i, t_i}(m_i)$ for each earthquake in the test set. We denote by $\ell_i$ the log-likelihood of the observed magnitude $m_i$ as predicted by MAGNET, 
\begin{align}
    \mathcal{L} &= -\langle \ell_i \rangle\ , 
    &
    \ell_i&=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)
    \label{eq:likelihood}
\end{align}
 $\mathcal{L}$ is the average minus log-likelihood, which the model is trained to minimize.  The values of $\ell_i$ for MAGNET and the leading benchmark are presented in Extended Material (EM) Fig. \ref{fig:labels_likelihood}.

@@

In addition to comparing the log-likelihood of the observed data, one can convert the prediction problem to a binary question: will the earthquake at the spacetime query $\textbf{x}_i, t_i$ have a higher magnitude than a threshold $m_t$? This can be easily calculated from the prediction, as the probability that $m_i>m_t$ is simply $\int_{m_t}^{\infty}p(m')dm'$.
This task, asking whether a given event will be a large earthquake, is more interpretable and can be evaluated using standard binary classification metrics such as area under the receiver operator (ROC) curve \cite{Murphy} and under the interpolated precision-recall (PR) curve \cite{buttcher_information_2010} which suit better an imbalanced data set such as ours. Fig.~\ref{fig:metrics}d-f and g-i show these metrics for all three regions examined and compare against the GR-variant models. Again, MAGNET outperforms all benchmarks.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/combined_barplots.pdf}
    \caption{
        \textbf{Benchmarking MAGNET. a}, \textbf{b}, \textbf{c}, Minus mean information content of our MAGNET model (red) and other common benchmark magnitude predictors (see labels on bars in the figure). A lower score indicates a better preforming model. Scores conditioned above the temporal incompleteness are marked by the horizontal dashed line on each corresponding bar. \textbf{d, e, f}, Receiver Operating Characteristic (ROC) and \textbf{g, h, i} the interpolated precision-recall (PR) curve for a binary classifier determining the next event will be large ($m>=4$). The performance of such a classifier can be quantified by the area under the curve (AUC), noted in each frame, color coded identically to the bar plots. For the AUC metrics, a higher score indicates a better preforming classifier.
        }
        \label{fig:metrics}
\end{figure}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/info_over_time_hauksson.pdf}
    \caption{
        \textbf{Information gain in MAGNET. a}, Information gain of individual events in the Southern California test set. Scattered dots indicate magnitude at event index, colored by the information gain per event (scale is indicated by colorbar), here warm is an advantage for MAGNET, while cool is an advantage for the benchmark. Secondary horizontal axis (grey) indicates the corresponding origin time. Cumulative Information gain (CIG) (dashed blue), temporally conditioned CIG (dashed yellow) and spatially conditioned CIG (dashed red) curves are superimposed on the scatter, demonstrating a constantly increasing function. \textbf{b-d}, Information gain (dashed blue) and temporally and spatially conditioned CIG (dashed yellow and red respectively) curves for Southern California, New Zealand and Japan test sets, respectively, all three regions examined in this study. An exception from the increasing trend can be observed for the spatially conditioned CIG of New Zealand (red curve in c). \textbf{e-g} Spatial distribution of spatially conditioned information gain for all regions in study. Maps show the difference in mean information content per bin, model's advantage is indicated by warm colors, disadvantage by cool colors, as defined by the colorbar adjacent to f.
    }
    \label{fig:info_gain_over_time}
\end{figure}

@@

Our analysis reveals a persistent increase in information gain across the test set for all investigated regions (Fig. \ref{fig:info_gain_over_time}b-d). Specifically, we show an average information gain of $>0.1$ nats per earthquake in the same region in Japan where Ogata et.~al.~\cite{ogata_exploring_2018} found no information gain. Similar plots for the other studied regions are shown in EM Fig.~\ref{fig:nz_japan_info_gain}.

@@

One such artifact is ``short-time incompleteness'' (STI): the fact that immediately following a large event the signal of subsequent small events might be buried in the coda of the mainshock~\cite{kagan_short-term_2004, stockman_forecasting_2023}. Thus, immediately after a large event it is less likely that a small event will be included in the catalog, not because it did not occur, but rather because it was not detected. Indeed, our model consistently allocates less probability mass to low magnitudes shortly after large events (see the Tohoku aftershock in Fig.~\ref{fig:intro_fig}b), and some of the information gain may be attributed to this effect.

@@

The instantaneous completeness magnitude, $m_c(t)$, is defined as the minimal magnitude above which all earthquakes can be detected. If we choose $\tilde m \geq m_c$ then, by definition, the model cannot present an information gain due to the STI effect.

We thus take $\Tilde{m}=m_c(t)$, which we estimate dynamically using the maximum curvature method \cite{wiemer_minimum_2000} within a window of 150 past and 150 current and future events. The resulting $m_c(t)$ curves are presented in the EM Fig. \ref{fig:temp_incompleteness}. The conditioned cumulative information gain (CIG) is then recalculated and presented as the yellow dashed curved in Fig.  \ref{fig:info_gain_over_time} for all three tested regions. 
Indeed, factoring out the temporal incompleteness reduces the information gain, though it is still significant: evidently, the conditioned CIG curves show a steadily increasing trend. The horizontal dashed lines on the bars in Fig \ref{fig:metrics}a-c indicate the mean conditioned scores, also presented in Table \ref{tab:mean_ll_all_benchmarks}. EM Table \ref{tab:mean_ll_benchmarks_main_text} displays the conditioned scores for all benchmarks tested in this study.

Similarly to the temporal domain, a spurious information gain may also be caused by temporally constant but spatially varying properties. For example, fault stresses, geometry and lithology may change in space resulting in a spatially-dependent GR distribution \cite{amitrano_brittle-ductile_2003, scholz_stress_2015, herrmann_revealing_2022, taroni_earthquake_2023}, and the local completeness magnitude is also generally spatially dependent. In such a case, the information gain of MAGNET does not imply that the magnitudes are correlated in time. 
To examine this, we preform two additional tests. First, we construct a spatially dependent GR distributions where $m_c$ is held constant ($=m_c^{\mbox{\footnotesize (train set)}}$). MAGNET outperforms the spatially varying GR benchmark for all regions in the unconditioned $\mathcal{L}$ score, as can be seen in EM Table \ref{tab:mean_ll_all_benchmarks}. The spatially varying GR shows an advantage over MAGNET for solely the New Zealand data set for the temporally conditioned $\mathcal{L}$. This advantage may arise from a the measurement artifact of spatial variation of $m_c$ originating from the properties of the seismic network.

In order to account for such spatially dependent artifacts we preform a simialr calculation. We once again condition MAGNET's results on a threshold according to Eq. \ref{eq:conditioned_likelihood}, this time choosing a spatially varying completeness $\Tilde{m}=m_c(x,y)$. We estimate $m_c(x,y)$ for a grid of coordinates spaced every $0.1^\circ$, for all events within a square centered at each grid point, with a side length of $0.2^\circ$.

@@

\section{Possible effects on operational earthquake forecasting and prediction}
Identifying precursory patterns in the seismic history can advance the effort of earthquake prediction a great deal\cite{mousavi_deep-learning_2022, mousavi_machine_2023, mignan_neural_2020, karimpouli_explainable_2023, bergen_machine_2019}. Nevertheless, discovering such signals with enough accuracy to prove useful for operational systems may still be further down the road.
However, an immediate application of the model may be a quick estimation of magnitude for early warning systems. Such a system does not require additional data sources or assumptions on the model as the timing and location of the earthquake to be estimated are known or estimated. 
Another immediate application is to incorporate such a history-dependent magnitude distribution into point-process rate forecasting models, such as the epidemic type aftershock sequence model (ETAS) or similar \cite{ogata_statistical_1988, ogata_statistics_2017, dascher-cousineau_using_2023}. As discussed above, such models currently assume no statistical dependence of the magnitude. Since larger earthquakes produce more aftershock, a more accurate estimation of magnitudes will have a positive compounding effect on the resulting statistics.

@@

Outputs of all encoders are concatenated and fed to a Fully Connected Neural Network (FCNN). The sizes of layers in the FCNN, as the number of layers themselves, are optimized by cross validation. A $\texttt{tanh}$ activation is used for the hidden layers, and a $\texttt{softplus}$ as an activation on the output layer.

@@

\subsection{Definition of benchmark models}
\subparagraph{Last $n$ events} The Gutenberg- Richter (GR) distribution fitted for the past $n$ events. This method follows Gulia \& Wiemer\cite{gulia_real-time_2019} for constructing a b-value time-series. If in the last $n$ events there are less than 10 earthquakes of magnitudes above $m_c^{train}$, then more events from the past are included, until this condition is met. $m_c^{train}$ is estimated using the maximal curvature method\cite{wiemer_minimum_2000}.

@@

\subparagraph{Spatially varying GR} Estimation of the local $b$ value is done by following the method of Taroni et.~al.~\cite{taroni_highdefinition_2021}. Seismicity data is 
binned into $0.1^\circ \times 0.1^\circ$ bins, and the GR distribution is fitted for each bin. $b$ for a specified location $x_i$ is calculated by the average of all $b$ values in bins within 30km around $x_i$. Estimates which consider less than 150 events are discarded.

@@

\newpage
\section*{Acknowledgements}
We thank Yehuda Ben-Zion, Assaf Inbal, Kelian Dascher-Cousineau, David Marsan, Eugenio Lippiello, Jiancang Zhuang and Yosihiko Ogata. for thoughtful discussions. YBS is supported by ISF grant 1907/22 and by Google Gift grant.

@@

\newpage
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/detailed_architecture.png}
    \caption{\textbf{MAGNET architecture.} Described in detail in the methods section.}
    \label{fig:architecture}
\end{figure}

\newpage
\begin{figure}[h!]
	\centering
        \includegraphics[width=1\textwidth]{figures/info_gain_scatter_NZ_and_Japan.png}
	\caption{
            \textbf{Information gain of MAGNET over the benchmark, New Zealand and Japan data sets.} Scattered dots indicate magnitude at event index, colored by the information gain per event. \textbf{a} Entire test set of New Zealand and a selected time span (\textbf{b}) focusing on the 2016 Te Araroa earthquake and the 2016 Kaikōura earthquake. \textbf{c} The entire Japan data set and a focused time span \textbf{d} of the 2011 Tōhoku earthquake and some of its aftershocks. 
         }
\label{fig:nz_japan_info_gain}
\end{figure}

\newpage
\begin{figure}[h!]
    \centering
        \includegraphics[width=1\textwidth]{figures/temporal_incompleteness.png}
    \caption{
    \textbf{Temporal Incompleteness for test sets of all regions examined.} Calculated temporal incompleteness for test sets is presented in red, superimposed over the events above (black, included in model evaluation) and below (grey) the completeness magnitude of the train set, $m_c^{(train)}$. Events are plotted by the index of their appearance in the test set. Data for Japan (c) is displayed only near the 2011 Tōhoku earthquake to reduce data density and enhance clarity of the figure.
    }
    \label{fig:temp_incompleteness}
\end{figure}

@@

\newpage
\begin{figure}[h!]
    \centering
    \includegraphics[width=1\textwidth]{figures/likelihood_scatter.png}
    \caption{
    \textbf{Log-likelihood of the true magnitude, $\ell_i$}. The likelihood for the true magnitude label of the test set for each data set, $\ell_i=\log\left(p_{\textbf{x}_i, t_i}(m_i)\right)$, as defined in main text. $\ell_i$ values of MAGNET and the last 300 events benchmark are shown in red and orange respectively. The blue dashed line indicates the values that the GR benchmark would obtain. In this representation, a higher value indicates a better score for an individual event. Both MAGNET and 300 last events benchmark to be well above the GR performance. It can be seen that for all regions MAGNET spreads the likelihoods wider, and specifically clearly outperforms the 300 last events benchmark in a significant portion of events. 
    }
    \label{fig:labels_likelihood}
\end{figure}
